### 2020-05-29 analysis/test.Rmd
Removing undeeded comparisons and more table colouring

---
title: "Supplementary Materials"
author: "Peter J. Kohler and Alasdair D. F. Clarke"
date:
output: bookdown::html_document2
---

I have set `echo=FALSE` so that most of the code chunks will not display. Please refer to the .Rmd file for source code.

```{r setup, include=FALSE}
library(bookdown)
library(tidyverse)
library(brms)
library(tidybayes)
library(ggridges)
library(latex2exp)
library(see)
library(ggthemes)
library(patchwork)
library(magick)

knitr::opts_chunk$set(echo = FALSE, message=FALSE, fig.align='center')

# specify some colours to use in the plots
v_cols <- ggthemes::ptol_pal()(6)

```

## Subgroup Information

Import subgroup information and display a table of the relationships that we will be investigating.

```{r}
source("scripts/import_subgroup_comparisons.R")

glimpse(subgroups)

opts <- options(knitr.kable.NA = "")

knitr::kable(index_table, caption = "subgroup index", escape = "F") %>% kableExtra::kable_styling(full_width = F)
```